Revisiting Multi-Domain Machine Translation

### 2020-10-23 2020/tacl2020/FinalVersion/main-2327-PhamMinhQuang.tex
Final version

\documentclass[11pt,a4paper]{article}
\usepackage{tacl2018v2}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{longtable}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage[draft]{todo}
\usepackage[normalem]{ulem}
\usepackage{xspace}
\usepackage{amsmath,amsfonts,amssymb}
%\setlength\titlebox{5cm}
%\colingfinalcopy % Uncomment this line for the final submission

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.
\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{red}{#1}}}
\newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{red}{#1}}}
\newcommand{\revision}[1]{#1}
\newcommand{\revisiondel}[1]{}
\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand{\system}[1]{\texttt{{#1}}}
\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
\newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
% \newcommand{\SB}[1]{\textcolor{green}{#1}}
% \newcommand{\SW}[1]{\textcolor{red}{#1}}
\newcommand{\SB}[1]{\textbf{#1}}
\newcommand{\SW}[1]{\underline{#1}}
\renewcommand\textfraction{.1}
\renewcommand\floatpagefraction{.95}
\newcommand{\sbcl}[2]{{\scriptsize #1 \hfill $|$ \hfill  #2}}
\usepackage{multirow}
\title{Revisiting Multi-Domain Machine Translation}
\fyFuture{A plea for more thoroughness in evaluation}

\author{MinhQuang Pham$^{\dag\ddag}$,\ \  Josep Crego$^{\dag}$,\ \  Fran\c cois Yvon$^{\ddag}$, \ \ Jean Senellart$^{\dag}$ \\ \\
  $^\ddag$Universit\'e Paris-Saclay, CNRS, LIMSI, 91400, Orsay, France \\
  \texttt{firstname.lastname@limsi.fr} \\
  $^\dag$SYSTRAN, 5 rue Feydeau, 75002 Paris, France \\
  \texttt{firstname.lastname@systrangroup.com} \\
}
\date{}

\begin{document}
\maketitle
\begin{abstract}
  When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training, and to robustly handle inputs from unexpected domains in testing. This multi-domain scenario has attracted a lot of recent work, that fall under the general umbrella of transfer learning.
In this study, we revisit multi-domain machine translation, with the aim to formulate the motivations for developing such systems and the associated expectations with respect to performance. Our experiments with a large sample of multi-domain systems show that most of these expectations are hardly met and suggest that further work is needed to better analyze the current behaviour of multi-domain systems and to make them fully hold their promises.
  \fyDone{Write abstract}
\end{abstract}

@@

%
% The following footnote without marker is needed for the camera-ready
% version of the paper.
% Comment out the instructions (first text) and uncomment the 8 lines
% under "final paper" for your variant of English.
% 
% \footnote{
%     %
%     % for review submission
%     %
%     \hspace{-0.65cm}  % space normally used by the marker
%     Place licence statement here for the camera-ready version. See
%     Section~\ref{licence} of the instructions for preparing a
%     manuscript.
%     %
%     % % final paper: en-uk version 
%     %
%     % \hspace{-0.65cm}  % space normally used by the marker
%     % This work is licensed under a Creative Commons 
%     % Attribution 4.0 International Licence.
%     % Licence details:
%     % \url{http://creativecommons.org/licenses/by/4.0/}.
%     % 
%     % % final paper: en-us version 
%     %
%     % \hspace{-0.65cm}  % space normally used by the marker
%     % This work is licensed under a Creative Commons 
%     % Attribution 4.0 International License.
%     % License details:
%     % \url{http://creativecommons.org/licenses/by/4.0/}.
% }

Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles. Given a training sample of matched source-target sentence pairs $(\src,\trg)$\jcDone{shoulnt trg be e as below ?} drawn from an underlying distribution $\mathcal{D}_s$, a model parameterized by $\theta$ (here, a translation function $h_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(h_\theta(\src), \trg)$. This approach ensures that the translation loss remains low when translating more sentences drawn from the same distribution.

Owing to the great variability of language data, this ideal situation is rarely met in practice, warranting the study of an alternative scenario, where the test distribution $\mathcal{D}_t$ differs from $\mathcal{D}_s$. In this setting, \emph{domain adaptation} (DA) methods are in order. DA has a long history in Machine Learning in general (e.g.\ \cite{Shimodaira00improving,Ben-David09atheory,Quinonero08dataset,Pan10asurvey})\fyDone{cite non covariate shift} and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}). Various techniques thus exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).
% In a nutshell, DA attempts to compensate for the distribution mismatch either by biasing the training sample to match ${\mathcal{D'}$ (eg.\ using resampling or instance-weighting strategies), or by adapting the

A seemingly related problem is \emph{multi-domain} (MD) machine translation \cite{Sajjad17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system is trained and tested with data from multiple domains. MD machine translation (MDMT) corresponds to a very common situation, where all available data, no matter its origin, is used to train a robust system that performs well for any kind of new input.
% As pointed out both by \newcite{Dredze08online} and \cite{Finkel09hierarchical}, this setting borrows much from DA, but also from multi-task learning \cite{Caruana97multitask,Yang15unified}.
If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT perform well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown? Should MDMT also be robust to new domains? How should it handle domain labeling errors? % , which makes their evaluation fragile

@@

In this paper, we formulate in a more precise fashion the requirements that an effective MDMT system should meet (Section~\ref{sec:requirements}). Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT systems and associated measurements to evaluate them (Section~\ref{sec:challenging}). In doing so, we also shed light on new problems that arise in this context, regarding for instance the accommodation of new domains in the course of training, or the computation of automatic domain tags.\fyDone{New problems - continuous learning, automatic domains} Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA. We show in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to the handling of a large amount of heterogeneous domains and to dynamically integrating new domains in training.\fyDone{Spell out conclusions}

\section{Requirements of multi-domain MT \label{sec:requirements}}
In this section, we recap the main reasons for considering a multi-domain scenario and discuss their implications in terms of performance evaluation.
\fyDone{Relation to theory, discussion of worst case scenario}
\fyDone{Insist on the work of Dredze}

@@

This definition seems to be the most accepted view of a multi-domain MT\footnote{An exception is \citep{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here. Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function. We now discuss the arguments that are put forward to proceed differently.
% --

% This is also the view that we adopt here, where we assume training inputs in $(\mathcal{X}\times{}\{1\dots{}D\})$, sampled from a mixture $\sum \lambda_{s,d}\mathcal{D}_{d} \times \{d\}$, and also consider various test scenarios, depending whether $\mathcal{D}_t$ is a fixed mixture of known domains, include new domain(s), in the supervised and unsupervised, and zero-shot settings.\fyTodo{explain zero shot}.

@@

A first motivation for moving away from the one-domain / one-system solution are practical  \cite{Sennrich13multidomain,Farajian17neural}: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop one single system instead of having to optimize and maintain multiple engines. The underlying assumption here is that the number of domains of interests can be large, a limiting scenario being fully personalized machine translation \cite{Michel2018extreme}.
% which is not always the case: for instance \cite{Britz17mixing,Zeng18multidomain,Jiang19multidomain} only consider 2-4 domains in their experiments.\fyTodo{Keep this last part ?}

@@

\fyDone{recap}
To recap, there are multiple arguments for adopting MDMT, some already used in DA settings, and some original. These arguments are not mutually exclusive; however each yields specific expectations with respect to the performance of this approach, and should also yield appropriate evaluation procedure. If the motivation is primarily computational, then a drop in MT quality with respect to multiple individual domains might be acceptable if compensated by the computational savings. If it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some under-resourced domains, over individually trained systems. If finally, it is to make the system more robust to unexpected or adversarial test distributions, then this is the setting that should be used to evaluate MDMT. The next section discusses ways in which these requirements of MDMT systems could be challenged.

% \cite[eq. (3)]{Wang20general} defines the loss of MDMT as the (macro) average of domain loss - all domains are equally important.
\section{Challenging multi-domain systems \label{sec:challenging}}
In this section, we propose seven\fyDone{check number} operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met. All these evaluations will rest on comparison of translation performance, and do not depend on the choice of a particular metric. To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best yardstick to assess subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}. Other important figures of merit for MDMT systems are the computational training cost and the total number of parameters.\fyDone{Insert discussion about scores}

\subsection{Multi-domain systems should be effective \label{ssec:effective}}
% \subsubsection{Performance}
A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. We thus derive the following requirements.

\paragraph{[P1-LAB]}\fyDone{Decide naming scheme}A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}. Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.\fyFuture{Try random label assignments ?} % Testing that MDMT actually benefit from domain tags is simple, notwithstanding the previous remarks above regarding evaluation procedures.

\paragraph{[P2-TUN]} Additionally, one can expect that MDMT will improve over fine-tuning \cite{Luong15stanford,Freitag16fast}, at least in domains where data is scarce, or in situations where several domains are close. To evaluate this, we perform two measurements, using a real as well as an artificial scenario. In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains. In the artificial scenario, we split a single domain in two parts which are considered as distinct in training. The expectation here is that a MDMT should yield a clear gain for both pseudo sub-domains, which should benefit from the supplementary amount of relevant training. In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.
\jcDone{I dont understand the prediction... why running MDMT on 2 artificial subdomains should show  gains over fine-tunning?}\fyDone{small loss with respect to non split ?}

@@

\paragraph{[P3-HET]}
The notion of a domain being a fragile one, an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ requirement [P2-TUN])\jcDone{what is P2.2?}, but also when intra-domain heterogeneity is hurting. This requirement is tested by artificially conjoining separate domains into one during training, hoping that the loss in performance with respect to the baseline (using correct domain tags) will remain small.

\paragraph{[P4-ERR]}
MDMTs should perform best when the true domain tag is known, but deteriorate gracefully in the face of tag errors; in this situation, catastrophic drops in performance are often observed\fyDone{cite \cite{McCloskey89catastrophic}}. This requirement can be assessed by translating test texts with erroneous domain tags and reporting the subsequent loss in performance.

\paragraph{[P5-UNK]}
A related situation occurs when the domain of a test document is unknown. Several situations need be considered: for domains seen in training, using automatically predicted domain labels should not be much worse than using the correct one. For test documents from unknown domains (zero-shot transfer), a good MD system should ideally outperform the default baseline that merges all available data.\fyDone{Systems react to unknown domains}
% These requirements are respectively labelled [P5.1] and [P5.2]\fyTodo{Not so sure of these ones}.

@@

We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}. \revision{We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}). Complementary experiments also use v12 of the News Commentary corpus (\domain{news}). Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}. To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%

We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training. Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}. We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value}

%for contrast experiments, we also use supplementary test sets from three other domains: the official Khresmoi testset \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, and IWSLT 2010 (Talk track) \cite{Paul10overview}. This enables us to evaluate the loss in performance when the test set is from a domain not seen in training.
% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.

@@

We measure the distance between domains using the $\mathcal{H}$-Divergence \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
%Formally, given test sets of size $m$ for domains $A$ and $B$, and $h(x)$ a trained domain predictor, $\mathcal{H}(A,B)$ is computed %as:
%$$
%\mathcal{H}(A,B) = 2(1 - [\frac{1}{m} \sum_{x:h(x) = B} \mathbb{I}( x \in A) + \frac{1}{m} \sum_{x: h(x) = A} \mathbb{I}(x \in B)]),
% $$
% where $\mathbb{I}$ is the indicator function.
Our discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model. We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values.\fyDone{Inform the classifier details}\fyDone{Insert tableau} Results in Table~\ref{tab:domaindist} show that all domains are well separated from all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.

@@

\begin{itemize}
\item a generic model trained on a concatenation of all corpora (\texttt{Mixed}). We develop two versions\footnote{In fact three: to enable a fair comparison with WDCMT, a RNN-based variant is also trained and evaluated. \revision{This system appears as \system{Mixed-Nat-RNN} in Table~\ref{tab:performance}}.} of this system, one where the domain unbalance reflects the distribution of our training data \revision{given in Table~\ref{tab:Corpora}} (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}). The former is the best option when the train mixture $\mathcal{D}^s$ is also expected in testing; the latter should be used when the test distribution is uniform across domains. Accordingly, we report two aggregate scores: a weighted average reflecting the training distribution, and an unweighted average, meaning that test domains are equally important.
%   As explained above, depending whether MDMT are expected to perform specially well on large domains, or to treat all domains equally, one or the other training distribution should be prefered.\jcDone{All systems cand be learned following both Bal and Nat. does Nat corresponds to wAVG in test? is wAVG defined somewhere?}
\item fine-tuned models \cite{Luong15stanford,Freitag16fast}\jcDone{i would use Luong and Manning 2015}, based on the \system{Mixed-Nat} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing. The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems adapted for one domain, with no parameter sharing across domains.
\revisiondel{We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a small adaptation module that is added (with residualconnections) on top of every Transformer layer (\system{FT-Res}).}
\end{itemize}

All models use embeddings and the hidden layers sizes of dimension~512. Transformers contain with 8 attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells. The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024. 
% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.\fyDone{Describe the block adaptation layer - voir slides} 
% All our systems are implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}.

>>>>

\documentclass[11pt,a4paper,acceptedWithA,copyedit]{article}
\usepackage{tacl2018v2}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{longtable}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage[draft]{todo}
\usepackage[normalem]{ulem}
\usepackage{xspace}
\usepackage{amsmath,amsfonts,amssymb}
%\setlength\titlebox{5cm}
%\colingfinalcopy % Uncomment this line for the final submission

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.
\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}

\newcommand{\revision}[1]{#1}
\newcommand{\revisiondel}[1]{}
\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand{\system}[1]{\texttt{{#1}}}
\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
\newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
% \newcommand{\SB}[1]{\textcolor{green}{#1}}
% \newcommand{\SW}[1]{\textcolor{red}{#1}}
\newcommand{\SB}[1]{\textbf{#1}}
\newcommand{\SW}[1]{\underline{#1}}
\renewcommand\textfraction{.1}
\renewcommand\floatpagefraction{.95}
\newcommand{\sbcl}[2]{{\scriptsize #1 \hfill $|$ \hfill  #2}}
\usepackage{multirow}

\title{Revisiting Multi-Domain Machine Translation}

\author{MinhQuang Pham$^{\dag\ddag}$,\ \  Josep Maria Crego$^{\dag}$,\ \  Fran\c cois Yvon$^{\ddag}$ \\ \\
  $^\ddag$Universit\'e Paris-Saclay, CNRS, LIMSI, 91400, Orsay, France \\
  \texttt{firstname.lastname@limsi.fr} \\
  $^\dag$SYSTRAN, 5 rue Feydeau, 75002 Paris, France \\
  \texttt{firstname.lastname@systrangroup.com} \\
}
\date{}

\begin{document}
\maketitle
\begin{abstract}
  When building machine translation systems, one often needs to make the best out of heterogeneous sets of parallel data in training, and to robustly handle inputs from unexpected domains in testing. This multi-domain scenario has attracted a lot of recent work, that fall under the general umbrella of transfer learning.
In this study, we revisit multi-domain machine translation, with the aim to formulate the motivations for developing such systems and the associated expectations with respect to performance. Our experiments with a large sample of multi-domain systems show that most of these expectations are hardly met and suggest that further work is needed to better analyze the current behaviour of multi-domain systems and to make them fully hold their promises.
\end{abstract}

@@

Data-based Machine Translation (MT), whether statistical or neural, rests on well understood machine learning principles. Given a training sample of matched source-target sentence pairs $(\src,\trg)$ drawn from an underlying distribution $\mathcal{D}_s$, a model parameterized by $\theta$ (here, a translation function $h_{\theta}$) is trained by minimizing the empirical expectation of a loss function $\ell(h_\theta(\src), \trg)$. This approach ensures that the translation loss remains low when translating more sentences drawn from the same distribution.

Owing to the great variability of language data, this ideal situation is rarely met in practice, warranting the study of an alternative scenario, where the test distribution $\mathcal{D}_t$ differs from $\mathcal{D}_s$. In this setting, \emph{domain adaptation} (DA) methods are in order. DA has a long history in Machine Learning in general (e.g.\ \cite{Shimodaira00improving,Ben-David09atheory,Quinonero08dataset,Pan10asurvey}) and in NLP in particular (eg.\ \cite{Daume06domain,Blitzer07domain,Jiang07instance}). Various techniques thus exist to handle both the situations where a (small) training sample drawn from $\mathcal{D}_t$ is available in training, or where only samples of source-side (or target-side) sentences are available (see \cite{Foster07mixture,Bertoldi09domain,Axelrod11domain} for proposals from the statistical MT era, or \cite{Chu18asurvey} for a recent survey of DA for Neural MT).

A seemingly related problem is \emph{multi-domain} (MD) machine translation \cite{Sajjad17neural,Farajian17multidomain,Kobus17domaincontrol,Zeng18multidomain,Pham19generic} where one single system is trained and tested with data from multiple domains. MD machine translation (MDMT) corresponds to a very common situation, where all available data, no matter its origin, is used to train a robust system that performs well for any kind of new input.
If the intuitions behind MDMT are quite simple, the exact specifications of MDMT systems are rarely spelled out: for instance, should MDMT perform well when the test data is distributed like the training data, when it is equally distributed across domains or when the test distribution is unknown? Should MDMT also be robust to new domains? How should it handle domain labeling errors?

@@

In this paper, we formulate in a more precise fashion the requirements that an effective MDMT system should meet (Section~\ref{sec:requirements}). Our first contribution is thus of methodological nature and consists of lists of expected properties of MDMT systems and associated measurements to evaluate them (Section~\ref{sec:challenging}). In doing so, we also shed light on new problems that arise in this context, regarding for instance the accommodation of new domains in the course of training, or the computation of automatic domain tags. Our second main contribution is experimental and consists in a thorough reanalysis of eight recent multi-domain approaches from the literature, including a variant of a model initially introduced for DA. We show in Section~\ref{sec:experiments} that existing approaches still fall short to match many of these requirements, notably with respect to the handling of a large amount of heterogeneous domains and to dynamically integrating new domains in training.

\section{Requirements of multi-domain MT \label{sec:requirements}}
In this section, we recap the main reasons for considering a multi-domain scenario and discuss their implications in terms of performance evaluation.

@@

This definition seems to be the most accepted view of a multi-domain MT\footnote{An exception is \citep{Farajian17multidomain}, where test translations rely on similarity scores between test and train sentences, rather than on domain labels.} and one that we also adopt here. Note that in the absence of further specification, the naive answer to the MD setting should be to estimate one translation function $\hat{h}_d(x)$ separately for each domain, then to translate using $\hat{h}(x,d) = \sum_{d'} h_{d'}(x) \indic{d' = d}$, where $\indic{x}$ is the indicator function. We now discuss the arguments that are put forward to proceed differently.

@@

A first motivation for moving away from the one-domain / one-system solution are practical  \cite{Sennrich13multidomain,Farajian17neural}: when faced with inputs that are potentially from multiple domains, it is easier and computationally cheaper to develop one single system instead of having to optimize and maintain multiple engines. The underlying assumption here is that the number of domains of interests can be large, a limiting scenario being fully personalized machine translation \cite{Michel2018extreme}.

@@

To recap, there are multiple arguments for adopting MDMT, some already used in DA settings, and some original. These arguments are not mutually exclusive; however each yields specific expectations with respect to the performance of this approach, and should also yield appropriate evaluation procedure. If the motivation is primarily computational, then a drop in MT quality with respect to multiple individual domains might be acceptable if compensated by the computational savings. If it is to improve statistical estimation, then the hope will be that MDMT will improve, at least for some under-resourced domains, over individually trained systems. If finally, it is to make the system more robust to unexpected or adversarial test distributions, then this is the setting that should be used to evaluate MDMT. The next section discusses ways in which these requirements of MDMT systems could be challenged.

\section{Challenging multi-domain systems \label{sec:challenging}}
In this section, we propose seven operational requirements that can be expected from an effective multi-domain system, and discuss ways to evaluate whether these requirements are actually met. All these evaluations will rest on comparison of translation performance, and do not depend on the choice of a particular metric. To make our results comparable with the literature, we will only use the BLEU score \cite{Papineni02bleu} in Section~\ref{sec:experiments}, noting it may not be the best yardstick to assess subtle improvements of lexical choices that are often associated with domain adapted systems \cite{Irvine13measuring}. Other important figures of merit for MDMT systems are the computational training cost and the total number of parameters.

\subsection{Multi-domain systems should be effective \label{ssec:effective}}
A first expectation is that MDMT systems should perform well in the face of mixed-domain test data. We thus derive the following requirements.

\paragraph{[P1-LAB]} A MDMT should perform better than the baseline which disregards domain labels, or reassigns them in a random fashion \cite{Joshi12multidomain}. Evaluating this requirement is a matter of a mere comparison, assuming the test distribution of domains is known: if all domains are equally important, performance averages can be reported; if they are not, weighted averages should be used instead.

\paragraph{[P2-TUN]} Additionally, one can expect that MDMT will improve over fine-tuning \cite{Luong15stanford,Freitag16fast}, at least in domains where data is scarce, or in situations where several domains are close. To evaluate this, we perform two measurements, using a real as well as an artificial scenario. In the real scenario, we simply compare the performance of MDMT and fine-tuning for domains of varying sizes, expecting a larger gain for smaller domains. In the artificial scenario, we split a single domain in two parts which are considered as distinct in training. The expectation here is that a MDMT should yield a clear gain for both pseudo sub-domains, which should benefit from the supplementary amount of relevant training. In this situation, MDMT should even outperform fine-tuning on either of the pseudo sub-domain.

@@

\paragraph{[P3-HET]}
The notion of a domain being a fragile one, an effective MDMT system should be able to discover not only when cross-domain sharing is useful (cf.\ requirement [P2-TUN]), but also when intra-domain heterogeneity is hurting. This requirement is tested by artificially conjoining separate domains into one during training, hoping that the loss in performance with respect to the baseline (using correct domain tags) will remain small.

\paragraph{[P4-ERR]}
MDMTs should perform best when the true domain tag is known, but deteriorate gracefully in the face of tag errors; in this situation, catastrophic drops in performance are often observed. This requirement can be assessed by translating test texts with erroneous domain tags and reporting the subsequent loss in performance.

\paragraph{[P5-UNK]}
A related situation occurs when the domain of a test document is unknown. Several situations need be considered: for domains seen in training, using automatically predicted domain labels should not be much worse than using the correct one. For test documents from unknown domains (zero-shot transfer), a good MD system should ideally outperform the default baseline that merges all available data.

@@

We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}. \revision{We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}}, the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; The JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}, documentations for KDE, Ubuntu, GNOME and PHP from Opus collection \cite{Tiedemann09news}, collectively merged in a \domain{it}-domain, Ted Talks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}). Complementary experiments also use v12 of the News Commentary corpus (\domain{news}). Most corpora are available from the Opus web site.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}. To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.

We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training.\footnote{The code for reproducing our train, dev and test datasets is available at \url{https://github.com/qmpham/experiments}.} Validation sets are used to chose the best model according to the average BLEU score \cite{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \cite{Neubig19compare-mt}. We report significant differences at the level of $p=0.05$.

@@

We measure the distance between domains using the $\mathcal{H}$-Divergence \cite{Ben-David09atheory}, which relates domain similarity to the test error of a domain discriminator: the larger the error, the closer the domains.
Our discriminator is a SVM independently trained for each pair of domains, with sentence representations derived via mean pooling from the source side representation of the generic Transformer model. We used the scikit-learn\footnote{\url{https://scikit-learn.org}} implementation with default values. Results in Table~\ref{tab:domaindist} show that all domains are well separated from all others, with \domain{rel} being the furthest apart, while \domain{talk} is slightly more central.

@@

\begin{itemize}
\item a generic model trained on a concatenation of all corpora (\texttt{Mixed}). We develop two versions\footnote{In fact three: to enable a fair comparison with WDCMT, a RNN-based variant is also trained and evaluated. \revision{This system appears as \system{Mixed-Nat-RNN} in Table~\ref{tab:performance}}.} of this system, one where the domain unbalance reflects the distribution of our training data \revision{given in Table~\ref{tab:Corpora}} (\system{Mixed-Nat}) and one where all domains are equally represented in training (\system{Mixed-Bal}). The former is the best option when the train mixture $\mathcal{D}^s$ is also expected in testing; the latter should be used when the test distribution is uniform across domains. Accordingly, we report two aggregate scores: a weighted average reflecting the training distribution, and an unweighted average, meaning that test domains are equally important.
\item fine-tuned models \cite{Luong15stanford,Freitag16fast}, based on the \system{Mixed-Nat} system, further trained on each domain for at most 20~000 iterations, with early stopping when the dev BLEU stops increasing. The full fine-tuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems adapted for one domain, with no parameter sharing across domains.
\revisiondel{We again contrast two versions: full fine-tuning (\system{FT-Full}), which may update all the parameters of the initial generic model; and the variant of \cite{Bapna19simple}, where fine-tuning only updates a small adaptation module that is added (with residualconnections) on top of every Transformer layer (\system{FT-Res}).}
\end{itemize}

All models use embeddings and the hidden layers sizes of dimension~512. Transformers contain with 8 attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells. The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024. 
Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers.

\subsection{Multi-domain systems \label{ssec:systems}}
